---
title: ""
# author: "Statistics for Engineers (MAT2001)- Lab"
author: "Introduction to R"
output:
  pdf_document: 
    toc: true
  html_document: 
    toc: true
    theme: cosmo
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```




# Introduction to R

-   R is a language and environment for statistical computing and
    graphics.
-   R provides a wide variety of statistical (linear and nonlinear
    modelling, classical statistical tests, time-series analysis,
    classification, clustering, …) and graphical techniques, and is
    highly extensible.
-   One of R’s strengths is the ease with which well-designed
    publication-quality plots can be produced, including mathematical
    symbols and formulae where needed.
-   R offers plenty of options for loading external data, including
    Excel, Minitab, SAS and SPSS files.
-   Here you can download [R](https://cloud.r-project.org/) and
    [RStudio](https://posit.co/download/rstudio-desktop/)

## Basics

-   After R is started, there is a console awaiting for input. At the
    prompt `>`, you can enter numbers and perform calculations.
-   The assignment operators are the left arrow with dash `<-` ,`->`and
    equal sign `=`.
-   The character `#` marks the beginning of a comment.All characters
    until the end of the line are ignored.

```{r echo=TRUE, message=TRUE}
# Example 
4+5

```

```{r echo=TRUE}
x <- 5
print(x)
```

## Create a vectors

*`c` function*

-   R functions are invoked by its name, followed by the parenthesis and
    arguments. The function `c` is used to combine three numeric values
    into a vector

The command `c(1,2,3,4,5)` combines the numbers 1,2,3,4 and 5 to a
vector.

```{r echo=TRUE}
c(1,2,3,4,5)

```

## Basic operations

R’s basic operators have the following precedence (listed in
highest-to-lowest order)

`^` exponentiation

`- +` unary minus and plus

`:` sequence operator

`%/% %%` integer division, remainder

`* /` multiplication, division

`+ -` addition, subtraction

```{r}
2^3^2
(2^3)^2
2^(3^2) 
sqrt(2)
```

## Seq operator and `seq` function

The expression $n_1:n_2$, generates the sequence of integers from $n_1$
to $n_2$

```{r}
# print the numbers 1 to 15
1:15
```

```{r}
# specifies interval and increment
seq(2,8,by=2)
```

```{r}
# specifies interval and the number of elements
seq(0,1,length=11)
```

Generating sequences of letters-lower case alphabets

```{r}
letters
letters[5:10]
```

### sequence of uppercase alphabets

```{r}
LETTERS
LETTERS[2:6]
```

Repeats-the command `rep`

```{r}
rep(2, times = 5)
rep(1:3, times = 4)
rep(1:3, each=2, times = 4)
```

## Matrix

-   Matrices are important objects in any calculation. A matrix is a
    rectangular array with p rows and n columns.

-   The parameter `nrow` defines the row number of a matrix.

-   The parameter `ncol` defines the column number of a matrix.

-   The parameter data assigns specified values to the matrix elements.

`matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE, dimnames = NULL)`

In R, a 4 × 2-matrix X can be created with a following command:

```{r}
x <- matrix(data = c(1:8), nrow = 4, ncol = 2)
print(x)

```

One can access a single element of a matrix with `x[i,j]`:

```{r}
x[3, 2]
```

We can get specific properties of a matrix:

```{r}
dim(x)
nrow(x)
ncol(x)

```

Assigning a specified number to all matrix elements:

```{r}
y <- matrix(data = 3, nrow = 2, ncol = 2)
print(y)
```

Construction of a diagonal matrix, here the identity matrix of a
dimension 2:

```{r}
diagonal <- diag(3, nrow = 2)
diagonal
```

### Transpose of a matrix X: X’

```{r}
z <- t(x)
z
```

### Basic matrix operation

```{r}
a <- matrix(1:9, nrow = 3, ncol = 3)

b <- matrix(11:19, nrow = 3, ncol = 3)

a+b
a-b
a/b
```

### Multiplication of a matrix with a constant

```{r}
# Note: x is already defined
x*5
```

Matrix multiplication: operator `%*%`

-   `crossprod()` computes $x^T×y$

Note: Command `crossprod()` executes the multiplication faster than the
conventional method with t(a)%\*%a

```{r}
a <- matrix(1:6, nrow = 3, ncol = 2)
a
b <- matrix(11:16, nrow = 2, ncol = 3)
b
a %*%b

crossprod(t(a), b)
```

### Access to rows, columns or submatrices

```{r}
x <- matrix(1:9, nrow = 3, ncol = 3)
x
x[2, ]
x[1:2,2:3]
```

## Sorting

`sort` function sorts the values of a vector in ascending order (by
default) or descending order.

```{r}
sort(c(20,50, 10, 30, 90,70, 80), decreasing = FALSE)
sort(c(20,50, 10, 30, 90,70, 80), decreasing = TRUE)
```

## Practice

1.  Create `a` sequence from 1 to 6
2.  Create `b` sequence from 7 to 12
3.  Create matrix for `a` and `b`.
4.  Calculate the basic operation for the two matrix.
5.  Arange the `b` sequence in descending order.

```{r, include=FALSE}
a <- matrix(seq(1, 6), nrow = 3)
b <- matrix(seq(7, 12), nrow = 3)
a+b
```

## Logical operator

```{r}
a <- 5
b <- 3

# Equal to
print(a == b) 

# Not equal to
print(a != b) 

# Greater than
print(a > b)  

# Less than or equal to
print(a <= b) 

# Less than
print(a < b)   
# Greater than or equal to
print(a >= b)  

```

```{r}
x <- c(1, 0, 3, 5)
y <- c(0, 3, 3, 2)

# Element-wise logical AND
print(x & y)  

# Element-wise logical OR
print(x | y)  

# Element-wise logical NOT on x
print(!x)  

# Comparison: x is greater than y
print(x > y) 

```

## Data Frames

-   In a data frame, we can combine variables of equal length,with each
    row in the data frame containing observations on the same unit.
-   Variables in a data frame may be numeric (numbers) or categorical
    (characters or factors).
-   Basic Syntax:
    `data_frame_name <- data.frame(column1 = vector1, column2 = vector2, ...)`

```{r}
id <- c(101, 102, 103)
name <- c("Alice", "Bob", "Charlie")
gender <- factor(c("Male", "Female", "Female"))
full_time <- c(TRUE, FALSE, TRUE)
salary <- c(50000, 60000, 55000)

# Creating the data frame
employee_data <- data.frame(ID = id, Name = name,Gender = gender,
                            FullTime = full_time, Salary = salary)

# Print the data frame
print(employee_data)

```

### Arithmetic mean

#### Ungrouped data

Find the mean value for 55,68,72,79,90,63,85,77,64,82, 55, 66, 89, 78,
67.

##### Manual method

```{r}
data <- c(55,68,72,79,90,63,85,77,64,82, 55, 66, 89, 78, 67)
data
avg <- sum(data)/length(data)
print(avg)
```

##### Using R command

```{r}
mean(data)
```

#### Grouped data - Discrete frequncy distribution

##### Manual method

```{r}
x <- table(data)
x
fx <- unique(data)*x
fx
N <- sum(x)
avg <- sum(fx)/N
avg
```

##### Using R command

```{r}
mean(data)
```

## Practice

1.  max(c(62,83,44,75)\^ -c(9,-3)) / min(c(52,62,71,85)\^c(2,3) ) -
    prod(c(1,2,1,2) \^ c(1,2))+max( c(12,13,14,15)\^c(2,3)) ?
2.  X1 \<- c(123,258,318,624), X2 \<- sqrt(X1 \^3) +X1/X1\^2-X1\*\*(1/2)
3.  X\<-matrix(nrow=3,ncol=3,data=
    c(10,20,30,40,50,60,70,80,90),byrow=F) then X[ ,2]?
4.  X\<-matrix(nrow=3, ncol=3, data = c(10,20,30,40,50,60,70,80,90) ,
    byrow=F) then X[2:3,2:3] ?
5.  sqrt(abs(seq(-6,6, by = 3)))?
6.  x \<- c(10, 75, 20, 35, 30, 40, 180, 50, 60, 27, 70, 67, 80, 50, 39,
    120) x[(x\>50)]? x[(x - 20 \> 40)] x[(x\^2 + 10 \> 50)]

